---
output: github_document
editor_options: 
  chunk_output_type: console
---

# easyRT

*Tools and examples for modelling Reaction Times in R*

**[WORK IN PROGRESS]**

```{r setup, echo = FALSE, warning=FALSE, message=FALSE}
options(digits = 3)
library(easyRT)

knitr::opts_chunk$set(
  collapse = TRUE,
  dpi = 450,
  fig.path = "man/figures/"
)

set.seed(111)
```


## Motivation 

This repo started as my attempt at understanding and implementing sequential models, starting with (Hierarchical) Drift Diffusion Models (DDMs) for reaction times in R. **Please don't hesitate** to open an issue to discuss and suggest things that could be improved or clarified.


## Content

Reaction time (RTs) have been traditionally modelled using traditional linear models (e.g., ANOVAs). However, it is problematic because RTs are **not** normally distributed. A popular mitigation method is to **transform** the data (e.g., usinga log-transform), but it is **not a good idea** ([Schramm & Rouder, 2019](https://doi.org/10.31234/osf.io/9ksa6)). Instead, one should use statistical models that **describe** or **generate** RT-like data.

You should start by reading:

- [**Lindelov's overview of RT models**](https://lindeloev.github.io/shiny-rt/): An absolute must-read.
- [**De Boeck & Jeon (2019)**](https://www.frontiersin.org/articles/10.3389/fpsyg.2019.00102/full): A paper providing an overview of RT models.

This repository contain the following vignettes:

- DDMs
- Ex-Gaussian

## Installation

```{r eval=FALSE}
remotes::install_github("DominiqueMakowski/easyRT")

library(easyRT)
```

## What does this package do?

Not much. It is mostly about its vignettes, but it also implement some convenience function to generate and plot drift diffusion models.

```{r warning=FALSE, message=FALSE}
sim <- ddm_data(drift = 1, bs = 1, bias = 0.5, ndt = 0.2)

ddm_plot(sim)
```

```

